EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-05-07 draft.tex
comments
Revises #5, still a few issues to check. @mariosky check comments

Developers can add new algorithm as stateless functions in
several ways and with two levels of integration. In the first level of
integration, the algorithm execution is not dependent on an initial
configuration created in the setup, and it only needs to take messages from the
input queue and return the modified population to the output queue. On the
second level, the algorithm receives a configuration from the initial setup,
specifying the initial algorithm parameters. We show an example of the configuration
file received by workers in Listing \ref{code:config}  in which we abbreviated
the population key. The message includes the current state
of the population, together with data about the problem, algorithm,
and its parameters.

@@

% Clarify this is the first level.
In the case of the first level of integration, a decoupled worker, developers
can implement a new stateless docker container following the description of a
worker container described in the previous section (\ref{workers}). The worker
container executes a demon that continuously pulls messages from the input queue
on the Redis host specified as an environmental variable {\texttt REDIS\_HOST}. If the new
population-based algorithm is written in python, developers can base the new
implementation on the provided worker container image, adding a new algorithm
function and worker and modifying the \texttt{main.py} file.
% say which function, what's its API, how can you test it - JJ
% this is not an API the need to write their own - Mario
% We can write an API later - M
If developers use an existing library for population-based algorithms, they need
to change the standard model of execution, that first creates a random
population because, in this case, we need to start with the population provided
as a parameter. Depending on the library, % Depending on which implementation? - JJ
it may be required to extract the population from the JSON message and replace
the population object created by the library. Again, the state of 
the population must be encoded as a JSON message to
be returned as a result. Additional data
must be returned in the JSON message, % how do you know the type of experiment? how do you add that? Which JSON key? - JJ
% I was wrong, it does not depend on the problem. - M
as depicted in  Listing \ref{code:result}. 
The output message contains the id of the worker container together
with records giving information about each iteration of the algorithm, 
including the best solution found. The controller uses this information 
to track the execution of the algorithm and keeping a log in the Redis 
container for posterior analyses.  % Is this mandatory? Optional? How is the algorithm going to use it? - JJ
% Mandatory and used for everything - M

\begin{lstlisting}[language=Python, caption=Fragment of output message, label=code:result ]
{
"time_stamp": 1588455316.698875,
"evals": [
  {
    "gen_num": 0, "best_fitness": -259.96925945939404,
    "best_solution": [ -4.9446848321611405, 
                        2.593007989886873, 
                        2.0011025031167673],
    "num_of_evals": 17
  },
  {
    "gen_num": 49, "best_fitness": -419.47555780783347,
    "best_solution": [ -0.9165294450108963, 
                        2.291966676472422, 
                        1.8527006568795485],
    "num_of_evals": 28
  }
],
"worker_id": "4f5eb12e-8278-40de-b0f3-4ac18171566d",
"message_counter": 1,
"message_id": "cbcd2e82-e232-4818-8aec-2cd072c7cfe9",
"best_score": -419.47555780783347
}
\end{lstlisting}

For the secondary level of integration, developers need to add code to the setup
container to specify additional configuration options. For instance,  % related to? What do you mean here?
in the current configuration message (see Listing \ref{code:config})
workers executing a GA, initialize the mutation and crossover 
probabilities randomly from a specific range of values. Users specify
this range on the experiment configuration.  % Not clear how you integrate this with the rest. An example would probably help. - JJ
% done - M
Migration between populations is independent of the type of algorithm, as the
controller treats messages as the same type of objects. But developers could
also change the controller to specify new rules of operation, for instance, to
change the type of migration depending on the algorithm. % Not clear how you would do this. Maybe a simple example would help - JJ
For example, in the case of swarm-based algorithms,  migration between  % Added - M
populations (swarms) can follow a topology, indicating valid connections 
between them (i.e., ring or hypercube). The controller
can apply these rules selectively only to swarm-based algorithms.

% This is the second level? - JJ
% either - M
On either level of integration, algorithm developers can upload the container
definition or images directly to a public repository like DockerHub. Another
option is to upload the container image definition (Dockerfile) to GitHub and
define a trigger to re-upload the image to the registry after an update. Having
images publicly available can contribute to the reproducibility of experiments.
Moreover, having more algorithms available can benefit researchers trying to
compare against other algorithms or increase the diversity of their algorithms
by adding more search methods. Finally, in some cloud services, having hosted 
images is a requirement, as it is the case of Amazon ECS.

% This is probably accessory. What difference would it make to upload it to GitHub?
% You also need to say in this section how it helps your overall objective. 
% Does it contribute to reproducibility? We've talked about that elsewhere. 
% Does it help diversity by making easy to add new algorithms? - JJ
% I explain it a bit more - M

@@

We aimed to learn if the proposed solution can efficiently
improve the scalability and performance of population-based optimization
algorithms. % If you want to prove it *improves*, you need to provide a baseline. What will be your baseline? You need to specify anyway in which way it really improves - JJ
Hence, in the following sections, we verify the following:

@@

\item Does having a multi-population enabled platform, with heterogeneous populations and
the support for mixing search strategies,  increases the performance of the
search by needing fewer function evaluations than a homogeneous setting? % But in this case the answer is no, right? - JJ

\item Can EvoSwarm as a cloud-native, container-based application,  be deployed
locally or in the cloud by just specifying the configuration of container images
and resources to an orchestration service?
\end{itemize}

To validate these questions, we used benchmark functions from the Continuous
Noiseless BBOB testbed, which is part of the Comparing Continuous Optimizers
(COCO) framework \cite{hansen2016coco}. The testbed includes 24 real-parameter,
single-objective benchmark functions, and the capability to provide additional
instances of each function. Each instance of a function has a different optimal
value. The standard benchmark of the testbed utilizes 15 instances per function
over 2, 3, 5, 10, 20, and 40 dimensions. The maximum number of Function
Evaluations (FEs) changes according to the dimension (D), and it is determined
by the expression $10^5 \cdot D$. As an example, if we have $D = 2$, the
maximum number of FEs is $200,000$.
% How do you manage to stop exactly at that number of evaluations in a distributed system? - JJ

>>>>

Developers can add new algorithm as stateless functions in
several ways and with two levels of integration. In the first level of
integration, the algorithm execution is not dependent on an initial
configuration created in the setup, and it only needs to take messages from the
input queue and return the modified population to the output queue. On the
second level, the algorithm receives a configuration from the initial setup,
specifying the initial algorithm parameters. We show an example of the configuration
file received by workers in Listing \ref{code:config}  in which we abbreviated
the population key (the ellipsis is used instead). The message includes the current state
of the population, together with data about the problem, algorithm,
and its parameters.

@@

In the case of the first level of integration, a decoupled worker, % a decoupler worker what? - JJ
developers
can implement a new stateless Docker container following the description in the previous section (\ref{workers}). The worker
container executes a daemon that continuously pulls messages from the input queue
on the Redis host specified as an environment variable {\texttt REDIS\_HOST}. If the new
population-based algorithm is written in Python, developers can base the new
implementation on the provided worker container image, adding a new algorithm
function and worker and modifying the \texttt{main.py} file.
% say which function, what's its API, how can you test it - JJ
% this is not an API the need to write their own - Mario
% We can write an API later - M
% Anyway you need to say what is the signature of the function. Also, if they can use the existing implementation, you need to specify the license - JJ
If developers use an existing library for population-based algorithms, they need
to change the standard model of execution, that first creates a random
population because, in this case, the algorithm need to start with the population provided
as a parameter. Depending on the library, % Depending on which library? - JJ
it may be required to extract the population from the JSON message and replace
the population object created by the library. Again, the state of 
the population must be encoded as a JSON message to
be returned as a result. Additional data
must be returned in the JSON message, 
as depicted in  Listing \ref{code:result}. 
The output message contains the id of the worker container together
with records giving information about each iteration of the algorithm, 
including the best solution found. The controller uses this information 
to track the execution of the algorithm and keep a log in the Redis 
container for posterior analyses.

\begin{lstlisting}[language=JSON, caption=Fragment of output message, label=code:result ]
{
"time_stamp": 1588455316.698875,
"evals": [
  {
    "gen_num": 0, "best_fitness": -259.96925945939404,
    "best_solution": [ -4.9446848321611405, 
                        2.593007989886873, 
                        2.0011025031167673],
    "num_of_evals": 17
  },
  {
    "gen_num": 49, "best_fitness": -419.47555780783347,
    "best_solution": [ -0.9165294450108963, 
                        2.291966676472422, 
                        1.8527006568795485],
    "num_of_evals": 28
  }
],
"worker_id": "4f5eb12e-8278-40de-b0f3-4ac18171566d",
"message_counter": 1,
"message_id": "cbcd2e82-e232-4818-8aec-2cd072c7cfe9",
"best_score": -419.47555780783347
}
\end{lstlisting}

For the secondary level of integration, developers need to add code to the setup
container to specify additional configuration options. For instance,  % related to? What do you mean here?
in the current configuration message (see Listing \ref{code:config})
workers executing a GA, initialize the mutation and crossover 
probabilities randomly from a specific range of values. Users specify
this range on the experiment configuration. 
Migration between populations is independent of the type of algorithm, as the
controller treats messages as the same type of objects. But developers could
also change the controller to specify new rules of operation, for instance, to
change the type of migration depending on the algorithm. 
For example, in the case of swarm-based algorithms,  migration between  
populations (swarms) can follow a topology, indicating valid connections 
between them (i.e., ring or hypercube). The controller
can apply these rules selectively only to swarm-based algorithms.

On either level of integration, algorithm developers can upload the container
definition or images directly to a public repository like DockerHub. Another
option is to upload the container image definition (Dockerfile) to GitHub and
define a trigger to re-upload the image to the registry after an update. Having
images publicly available can contribute to the reproducibility of experiments.
Moreover, having more algorithms available can benefit researchers trying to
compare against other algorithms or increase the diversity of their algorithms
by adding more search methods. Finally, in some cloud services, having hosted 
images is a requirement, as it is the case of Amazon ECS.

@@

We aimed to learn if the proposed solution can efficiently
improve the scalability and performance of population-based optimization
algorithms. % If you want to prove it *improves*, you need to provide a baseline. 
            %What will be your baseline? You need to specify anyway 
            %in which way it really improves - JJ
            % The baseline is a single population GA or PSO? - M
Hence, in the following sections, we verify the following:

@@

\item Does having a multi-population enabled platform, with heterogeneous populations and
the support for mixing search strategies,  increases the performance of the
search by needing fewer function evaluations than a homogeneous setting? % But in this case the answer is no, right? - JJ
                                                                         % Is yes, but is in the second experiment - M
\item Can EvoSwarm as a cloud-native, container-based application,  be deployed
locally or in the cloud by just specifying the configuration of container images
and resources to an orchestration service?
\end{itemize}

To validate these questions, we used benchmark functions from the Continuous
Noiseless BBOB testbed, which is part of the Comparing Continuous Optimizers
(COCO) framework \cite{hansen2016coco}. The testbed includes 24 real-parameter,
single-objective benchmark functions, and the capability to provide additional
instances of each function. Each instance of a function has a different optimal
value. The standard benchmark of the testbed utilizes 15 instances per function
over 2, 3, 5, 10, 20, and 40 dimensions. The maximum number of Function
Evaluations (FEs) changes according to the dimension (D), and it is determined
by the expression $10^5 \cdot D$. As an example, if we have $D = 2$, the
maximum number of FEs is $200,000$.
% How do you manage to stop exactly at that number of evaluations in a distributed system? - JJ
% In the controller we check the number of messages received, we assume that each message 
% needs a X amount of #FEs we stop at the maximum number of messages parameter.
% If the solution is found we stop sending messages and ignore messages from that
% experiment ID - M
